\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-05-14 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Add more text about memory satisfaction

With the future region relation in place, we define the future world relation as follows:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
            { W' \future W }
\end{mathpar}
The relation says that each of the three worlds must be an extension of the past world and for each of the existing regions that must be a future region.
% injective function
Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard:esop18} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.

@@

\subsubsection{Solving the recursive domain equation}
\label{subsubsec:rec-dom-eq}
%solving the recursive domain equation means that we can present the actual definitions
In the previous sections, we have sketched what we would like our worlds to look like.
However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
% Move to a different domain
Therefore, we need to move to a different domain with enough structure for such an equation to have a solution.
Specifically, we move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
Intuitively, a c.o.f.e. is a set with some extra structure.
Specifically, c.o.f.e.'s have a step-index which provide sufficient structure for us to solve recursive domain equations.
In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.

% Cofe's
In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
We loosely follow the presentation in \citet{Birkedal:tutorial-notes}.
\begin{definition}[Ordered family of equivalences (o.f.e.)]
An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and for each $n$ $\nequal$ is an equivalence relation.
The pair must satisfy the following properties
\begin{enumerate}
\item $\nequal[0]$ is the total relation.
\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
\end{enumerate}
\end{definition}
\noindent One can think of the indexes on the family of equivalences as a measure of precision.
The larger $n$ is, the more refined the equivalence may be.
On the other hand as $n$ decreases, the equivalence becomes more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
At index 0, all precision is lost and the equivalence cannot distinguish anything.
% TODO: Lau write intuition for sequences and why they are necessary.
\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
  \label{def:cauchy-sequence}
  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
  in $X$.

@@

  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
  \end{align*}
\end{definition}
Functions between c.o.f.e.'s must retain the added structure relatively to sets.
To this end, we require all our functions to be \emph{non-expansive} which means that $n$-equivalences are preserved.
A function is called \emph{contractive} when it not only retain equivalences but makes them more precise.
\begin{definition}
  \label{def:nonexpansive-contractive-ofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
  We 
  The function $f$ is
  \begin{description}
  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal_Y f(x')
\]
  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
  \qedhere
\]
  \end{description}
\end{definition}
\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
However, this is not the only structure we would like to impose on our sets.
In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
The future world relations is essentially a preorder structure that we would like to impose on our sets.
We impose this structure on our sets by moving to \emph{preordered c.o.f.e.'s}.
% preordered cofe
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The new structure we add to the set, the preorder, must be consistent with the existing structure which is why it must preserve limits.

Functions between preordered c.o.f.e's must preserve the ordering which means that they must be monotone with respect to the ordering.
% monotone functions
\begin{lemma}
  \label{def:monotone-preordered-cofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty},\future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} ,\future_Y\right)$ be two preordered c.o.f.e.'s and $f$ a function from the set $X$ to the set $Y$.
  We 
  The function $f$ is
  \begin{description}
  \item[monotone] when for all $x, x' \in X$,
\[
  x' \future_X x \implies f(x) \future_Y f(x')
\]
  \end{description}
\end{lemma}
% Not sure this is necessary:
% % preordered cofe constructions
% %% cofe to preordered cofe
% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
% \end{lemma}
In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.

@@

Next, we define the preordered c.o.f.e.\ of monotone and non-expansive functions.
%% Monotone, non-expansive
\begin{lemma}[Preordered c.o.f.e. of monotone and non-expansive functions]
  Given preordered c.o.f.e.\ $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and a c.o.f.e.\ $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows 
  \[
    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
  \]
  and preorder
  \[
    h \future h' \text{ iff } \forall r \in \dom(h) \ldotp h(w) = h'(w)
  \]
\end{lemma}
Next, we define the product preordered c.o.f.e.
%% Product
\begin{lemma}[Product preordered c.o.f.e.]
  Two preordered c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \future_Y \right)$ can be composed to a product preordered c.o.f.e $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
  where the family of equivalences are defined as follows
  \[
    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
  and the order as follows
  \[
    (x,y) \future (x',y') \text{ iff } x \future_X x' \wedge y \future_Y y'
  \]
\end{lemma}
We use the product preordered c.o.f.e.\ when we just need to compose two existing preordered c.o.f.e.'s into a new preordered c.o.f.e.
However, in the case of our regions, this do not suffice as they need to use the future region relation we defined in Section~\ref{subsubsec:ft-and-revocation}.
We therefore define separate preordered c.o.f.e.'s that we use to define these.
\begin{lemma}[$\Regions$ preordered c.o.f.e.]
  Given preordered c.o.f.e.\ $\left(H, \left(\nequal_H\right)_{n=0}^{\infty}, \future_H \right)$ define $\left(\{\spatial\} \times H + \{\spatialo\} \times H + \revoked, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$
  where $\future$ is the future region relation from Section~\ref{subsubsec:ft-and-revocation} and
  \[
    r \nequal r' \text{ iff } n=0 \vee \left\{
      \begin{aligned}
      (\_,h),(\_,h') \in \{ \spatial  \} \times H \wedge h \nequal h' \vee \\
      (\_,h),(\_,h') \in \{ \spatialo \} \times H \wedge h \nequal h' \vee \\
      r,r' = \{revoked\}      
    \end{aligned}\right.
  \]
\end{lemma}

%% Sum
Sum c.o.f.e.'s are defined in a similar manner.

\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
%% Partial function from set
\begin{lemma}[Preordered c.o.f.e.\ of partial functions]
  Given a set $S$ and a preorder c.o.f.e $P$, the set of partial functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
  \[
    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
  \]
  and preorder
  \[
    f \future g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \future g(x)
  \]
\end{lemma}

%% Later cofe
\begin{lemma}[$\blater$ preordered c.o.f.e.]
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right) = \left(X, \left(\nequal[n]'\right)_{n=0}^{\infty}, \future \right)
  \]
  where
  \[
    x \nequal[0]' x' \wedge x \nequal[n]' x' \texttt{ iff } x \nequal[n+1] x'
  \]
\end{lemma}

% partial functions
%

% New world definitions
The world definitions we presented so far needs to move to the domain of preordered c.o.f.e's. Morally, the definitions are the same, though. The definition of the two kinds of regions now uses $\URel{}$ rather than $\Rel$, and the world indexed-functions are monotone. Finally, all functions are now non-expansive.
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
\end{multline*}
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
    \{\revoked\}
  \end{array} \right.
\]

% Reference to solution?
%% Aleš note? Tutorial note?
%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?

% Argue, informally, that it is okay

% Solution theorem

\subsubsection{Joining worlds}

In order to express this, we will need to split the ownership of the world in disjoint parts.
%% Disjoint union \oplus
To this end, we introduce the operator $\oplus$ which is a disjoint union.
\begin{definition}[$\oplus$, disjoint union of ownership]
  \[
    W_1 \oplus W_2 = W
  \text{ iff }
  \begin{array}[t]{l}
    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
  \end{array}
  \]
\end{definition}

@@

\subsubsection{Memory satisfaction}
\label{subsubsec:mem-sat}
% Well-formedness of world
% specification satisfaction
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \text{ is defined} \wedge\\
      W = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}} \wedge\\
      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}, \ms_{T,f}, \ms_{S,f}, \ms_S',\overline{\sigma}\ldotp \\
      \quad \ms_S = \ms_{f,S} \uplus \ms_S' \wedge \\
      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \uplus \ms_{T,f} \wedge \\
      \quad \dom(\ms_{\var{T,stack}} \uplus \var{T,free\_stack}) = [\baddr_\stk,\eaddr_\stk] \wedge \\
      \quad \{\baddr_\stk -1,\eaddr_\stk + 1\} \in \dom(\ms_{T,f}) \wedge \\
      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
      \quad \npair{(\overline{\sigma},\ms_S',\ms_\var{T,heap})} \in \lrheap(\pwheap)(W_{\var{heap}})
    \end{array}
  \right.
\]

\[
  \memSatStack{\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      W_\var{stack} = \pwpriv \wedge \\
      \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
      \forall i \in \{0,\dots,m\} \ldotp (\dom(\ms_i) \neq \emptyset \wedge\\
      \quad \forall i < j \ldotp \forall a \in \dom(\ms_i) \ldotp \forall a' \in \dom(\ms_j) \ldotp \stkb < a < a') \wedge\\
      \exists R_\ms : \dom(\activeReg{W_\var{stack}}) \fun \MemSeg \times \Addr \times \MemSeg \ldotp \\
      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_3(R_\ms(r)) \wedge \\
      \quad \ms_0 \uplus \dots \uplus \ms_m = \biguplus_{r \in \dom(\activeReg{W_\var{stack}})} \pi_1(R_\ms(r)) \wedge \\
      \quad \exists R_W : \dom(\activeReg{W_\var{stack}}) \fun \World \ldotp \\
      \qquad W = \bigoplus_{r \in \dom(\activeReg{W_\var{stack}})} R_W(r) \wedge \\
      \qquad \forall r \in \dom(\activeReg{W_\var{stack}}), n' < n\ldotp \\
      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(r))} \in W_\var{stack}(r).H \; \xi^{-1}(R_W(r)) \wedge\\
      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc \wedge\\
      \qquad \quad \exists i \ldotp \opc_i = W_\var{stack}(r).\opc \wedge \ms_i = \pi_1 (R_\ms(r))
    \end{array}
  \right.
\]

>>>>

With the future region relation in place, we define the future world relation as follows:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \future W.i(r) }
            { W' \future W }
\end{mathpar}
The relation says that each of the three worlds must be an extension of the past world and for each of the existing regions that must be a future region.
% injective function
Note that the future world relation has a mapping function $m_i$ which allows us to change the naming of regions in future worlds.
The definition is a generalization of the standard definition where $m_i$ would be the identity\footnote{In \citet{Skorstengaard_reasoning_2017} the future region relation and the reasoning about the awkward example could have been simplified with this future world relation.}.

@@

\subsubsection{Solving the recursive domain equation}
\label{subsubsec:rec-dom-eq}
%solving the recursive domain equation means that we can present the actual definitions
In the previous sections, we have sketched what we would like our worlds to look like.
However, the worlds we want constitute a self-referential domain equation for which no solution exists in set theory.
% Move to a different domain
Therefore, we need to move to a different domain with enough structure for a solution to exist for such an equation.
We move to a setting where instead of sets we have c.o.f.e.'s (complete ordered families of equivalences), instead of functions we have non-expansive functions, and instead of relations we have downwards-closed relations.
Intuitively, a c.o.f.e. is a set with some extra structure.
Specifically, c.o.f.e.'s have a step-index which provide sufficient structure for us to solve recursive domain equations.
In this section we present the relevant definitions, the final world domain equation, and sketch what needs to be proven to get a solution to the recursive domain equation.

% Cofe's
In the following, we present complete ordered family of equivalences \citep{di_gianantonio_2002}.
We loosely follow the presentation in \citet{birkedal_taste_2014} and introduce the necessary definitions needed for this particular setting.
\begin{definition}[Ordered family of equivalences (o.f.e.)]
An \emph{ordered family of equivalences} is a pair $\cofe{X}$ where $X$ is a set and for each $n$ $\nequal$ is an equivalence relation.
The pair must satisfy the following properties
\begin{enumerate}
\item $\nequal[0]$ is the total relation.
\item For all $n \in \nats$, $\nequal[n] \supseteq \nequal[n+1]$.
\item For all $x,y \in X$,  if for all $n \in \nats$ $x \nequal y$, then $x = y$.
\end{enumerate}
\end{definition}
\noindent One can think of the indexes on the family of equivalences as a measure of precision.
The larger $n$ is, the more refined the equivalence may be.
On the other hand as $n$ decreases, the equivalence becomes more and more imprecise and may be unable to distinguish elements that where distinguishable at higher indices.
At index 0, all precision is lost and the equivalence cannot distinguish anything.
% TODO: Lau write intuition for sequences and why they are necessary.
\begin{definition}[Complete ordered family of equivalences (c.o.f.e.)]
  \label{def:cauchy-sequence}
  A \emph{complete ordered family of equivalences} is an ordered family of equivalences
  $\cofe{X}$ such that every Cauchy sequence in $X$ has a limit
  in $X$.

@@

  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
  \end{align*}
\end{definition}
% TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
Functions between c.o.f.e.'s must retain the added structure relatively to sets.
To this end, we require all our functions to be \emph{non-expansive} which means that $n$-equivalences are preserved.
A function is called \emph{contractive} when it not only retain equivalences but makes them more precise.
\begin{definition}
  \label{def:nonexpansive-contractive-ofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
  We 
  The function $f$ is
  \begin{description}
  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal_Y f(x')
\]
  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
  \qedhere
\]
  \end{description}
\end{definition}
\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
However, this is not the only structure we would like to impose on our sets.

% Not sure this is necessary:
% % preordered cofe constructions
% %% cofe to preordered cofe
% \begin{lemma}[c.o.f.e.'s are preordered c.o.f.e's]
%   A c.o.f.e $\cofe{X}$, is also a preordered c.o.f.e. with the order $=$.
% \end{lemma}
In order to express the domain equation, we must lift all parts of the world definition from sets to c.o.f.e.'s.
That is, we need to lift the sets of partial functions, monotone non-expansive functions, and relations, respectively to c.o.f.e.'s.

@@

Next, we define the c.o.f.e.\ of monotone and non-expansive functions.
%% Monotone, non-expansive
\begin{lemma}[The c.o.f.e. of monotone and non-expansive functions]
  Given c.o.f.e.'s $\left(W,\left(\nequal[n]_W\right)_{n=0}^{\infty}, \future_W\right)$ and $\left(U,\left(\nequal[n]_U\right)_{n=0}^{\infty}\right)$ and a preorder $\future$ on $W$, define the preordered c.o.f.e.\ $\left(W \monnefun U,\left(\nequal[n]\right)_{n=0}^{\infty}, \future \right)$ as follows 
  \[
    h \nequal h' \text{ iff } \forall w \in \dom(h) \ldotp h(w) \nequal_U h'(w)
  \]
\end{lemma}
Next, we define the product c.o.f.e.
%% Product
\begin{lemma}[Product c.o.f.e.]
  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
  where the family of equivalences are defined as follows
  \[
    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
\end{lemma}
Similarly to the product c.o.f.e. we have a sum c.o.f.e.
%% Sum
\begin{lemma}[Sum c.o.f.e.]
  Given two c.o.f.e's $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty}, \right)$ define the sum c.o.f.e.\ as $\left(X \cup Y, \left(\nequal\right)_{n=0}^{\infty}, \right)$
  where the family of equivalences are defined as follows
  \[
    a \nequal b \text{ iff }
    \left\{
      \begin{aligned}
        a,b \in X \wedge a \nequal_X b \vee \\
        a,b \in Y \wedge a \nequal_Y b
      \end{aligned}
    \right.
  \]
\end{lemma}
\todo[inline]{Lau: Do we need to use the future world relation here? Generally where do we need to use the prior defined relations?}
%% Partial function from set
\begin{lemma}[The c.o.f.e.\ of partial functions]
  Given a set $S$ and a c.o.f.e $P$, the set of partial non-expansive functions $S \parfun P$ can be lifted to a preordered c.o.f.e.\ with family of equivalences
  \[
    f \nequal[0] g \wedge f \nequal g \text{ iff } \forall x \in \dom(f) \ldotp f(x) \nequal g(x)
  \]
\end{lemma}

In Section~\ref{subsubsec:ft-and-revocation}, we described a future world relation which models changes to the memory.
The future world relations is essentially a preorder structure that we would like to impose on our sets.
We impose this structure on our sets by moving to \emph{preordered c.o.f.e.'s}.
% preordered cofe
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The new structure we add to the set, the preorder, must be consistent with the existing structure which is why it must preserve limits.

Functions between preordered c.o.f.e's must preserve the ordering which means that they must be monotone with respect to the ordering.
% monotone functions
\begin{lemma}
  \label{def:monotone-preordered-cofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty},\future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} ,\future_Y\right)$ be two preordered c.o.f.e.'s and $f$ a function from the set $X$ to the set $Y$.
  We 
  The function $f$ is
  \begin{description}
  \item[monotone] when for all $x, x' \in X$,
\[
  x' \future_X x \implies f(x) \future_Y f(x')
\]
  \end{description}
\end{lemma}
% TODO a preordered c.o.f.e. w/ our world?

In some sense, the step-index in a c.o.f.e. gives us a metric that we may use to find a fixed-point.
However, for a fixed-point to exist, the metric must decrease before a self-reference.
As this is not an inherent property of c.o.f.e.'s, we have a c.o.f.e.\ construction called \emph{later} with the sole purpose of decrementing the index.
%% Later cofe
\begin{lemma}[$\blater$ preordered c.o.f.e.]
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right) = \left(X, \left(\nequal[n]'\right)_{n=0}^{\infty}, \future \right)
  \]
  where
  \[
    x \nequal[0]' x' \wedge x \nequal[n]' x' \text{ iff } x \nequal[n+1] x'
  \]
\end{lemma}
The later needs to guard all negative self-references in a recursive domain equation in order for a solution to exist.
The modality does not have to be placed directly on the self-reference.
For instance, in the following it is placed on the world itself.
The later placement does, however, affect what the logical relation definitions in Section~\ref{subsec:logical-relation} looks like, so in the end it is a matter of taste.
% Solution theorem
\begin{theorem}
  \label{thm:rec-dom-eq-sol}
  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
  \[
    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
  \]
  and for $\hat{W},\hat{W}' \in \Wor$
  \[
    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
  \]
  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
\[
  \Worldh = \RegionName \parfun (\Regions + \Regionh)
\]
and
\[
  \Worlds = \RegionName \parfun (\Regions \times \Addr)
\]
and
\[
  \Worldfs = \RegionName \parfun \Regions
\]
where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps^2})
\end{multline*}
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\uplus \\ 
    \{\revoked\}
  \end{array} \right.
\]
\end{theorem}
% TODO informally describe how solution to recursive domain equation is found (in terms of what we have to show, not the fixed-point theorem).
%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
%% Aleš note? Tutorial note?

The domains in Theorem~\ref{thm:rec-dom-eq-sol} are made from the c.o.f.e.\ and preorder c.o.f.e.\ constructions we presented earlier in this section.
They are also the definitions we use in our world definition.
We do not define $\Wor$ to be our worlds.
Instead, we define our world as
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]
which is easier as we only have to use $\xi$ and move under a later when we have gone into the region and need to work with $\Wor$.

\subsubsection{Joining worlds}
% TODO rewrite this introduction
In order to express this, we will need to split the ownership of the world in disjoint parts.
%% Disjoint union \oplus
To this end, we introduce the operator $\oplus$ which is a disjoint union.
\begin{definition}[$\oplus$, disjoint union of ownership]
  \[
    W_1 \oplus W_2 = W
  \text{ iff }
  \begin{array}[t]{l}
    \dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand \\
    \dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand \\
    \dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand \\
    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
  \end{array}
  \]
\end{definition}

@@

\subsubsection{Memory satisfaction}
\label{subsubsec:mem-sat}
% Specification
The world can be seen as a specification of the memory contents.
This means that we need to define what it means for a pair of \trgcm{} and \srccm{} memories to satisfy the specification.
% Call stack
The world also keeps track of the structure of the call stack, the allowed uses of designated seals, and linear capability authority, so these things also influence memory satisfaction.
% Well-formedness
The memory satisfaction also acts as a well-formedness judgement.
Our world definition allows regions to be overlapping in the sense that two regions may allow memory segments with the same addresses.
We want the regions to be disjoint, so we easily can get the invariants of each memory segment.

Memory satisfaction is split into four definitions.
At the top-level we have $\memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W}$ which relates the source memory, $\ms_S$, $\ms_\stk$, and $\stk$, from \srccm{} to the target memory $\ms_T$ from \trgcm{}.
The top memory satisfaction splits the target memory in three parts, one for each of the three kinds of source memory.
It also splits the world in three to distribute the authority of the world.
The three parts of $\ms_T$ are related to $\ms_S$, $\ms_\stk$, and $\stk$ by the relations $\lrheap$, $\lrstk[]$, and $\lrfree[]$, respectively. 
The $\lrheap$ relation relates the \srccm{} heap (non-stack memory) to the corresponding heap memory on \trgcm{}.
The $\lrheap$ relation also ensures that the seal invariants associated with each regions do not put invariants on the same seals.
The $\lrstk[]$ relation relates the stack of encapsulated stack frames on \srccm{} to the corresponding memory on \trgcm{}.
The layout of the stack determines the call order, so $\lrstk[]$ also makes sure that the stack frames have the correct order relative to each other as well as the base address of the stack.
Finally, $\lrfree[]$ relates the free part of the stack of \srccm{} to the corresponding memory segment on \trgcm{}.
$\lrfree[]$ also makes sure that that base address of the stack is actually part of the stack.
\begin{definition}[Heap relation]
For a set of seals $\overline{\sigma}$, memory segments $ms$ and $\ms_T$, and worlds $W$ and $W'$, we define the heap relation $\lrheap$ as:
\[
  \npair{(\overline{\sigma},ms,\ms_T)} \in \lrheap(\pwheap)(W') = 
  \left\{
    \begin{array}{l}
      \exists R_\ms : \dom(\activeReg{\pwheap}) \fun \MemSeg \times \MemSeg \wedge \\
      \quad \ms_T = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_2(R_\ms(r)) \wedge \\
      \quad \ms = \biguplus_{r \in \dom(\activeReg{\pwheap})} \pi_1(R_\ms(r)) \wedge \\
      \quad \exists R_W : \dom(\activeReg{\pwheap}) \fun \World\ldotp\\
      \qquad W' = \oplus_{r \in \dom(\activeReg{\pwheap})} R_W(r) \wedge\\
      \qquad \forall r \in \dom(\activeReg{\pwheap}), n' < n \ldotp \\
      \qquad \quad \npair[n']{R_\ms(r)} \in  \pwheap(r).H \; \xi^{-1}(R_W(r)) \wedge\\
      \exists R_\var{seal} : \dom(\activeReg{\pwheap}) \fun \powerset{\Seal} \wedge\\
      \quad \biguplus_{r \in \dom(\activeReg{\pwheap})} R_\var{seal}(r)) \subseteq \overline{\sigma} \wedge\\
      \quad \forall r \in \dom(\activeReg{\pwheap})\ldotp \dom(\pwheap(r).H_\sigma) = R_\var{seal}(r)
    \end{array}
  \right.
\]
\end{definition}
Memory satisfaction, and thus also the heap relation, only considers the non-$\revoked$ regions.
The $\lrheap$-relation uses the function $\activeReg{}$ to erase all the $\revoked$ regions from the world.

To a large extend, the way $\lrheap$ is defined is pretty standard.
$\lrheap$ assumes the existence of a partitioning of the \trgcm{} and \srccm{} heap memories that can be turned into pairs each satisfying the invariant of a region.
The heap satisfaction must also respect the world as an authority specification, so the heap satisfaction also partition the authority of the world.
Each of the memory segment pairs must be in the region invariant with respect to a specific world partition which makes sure that uniqueness of linearity of capabilities is respected.
